---
title: "arima?"
output: html_document
date: "2025-04-15"
---

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(lubridate)
library(parallel)
library(forecast)
library(ggplot2)
library(tidyr)


data <- read_csv("NE_LINCOLN_2008_2024.csv")

lincoln25 <- read_csv("NE_LINCOLN_2025 (1).csv")
data$time <- as.POSIXct(data$time,format = "%m/%d/%Y %H:%M:%S")



```

Decomposition, arima model, forecast
```{r}
data <- data %>%
  drop_na(gamma2)

plot.ts(data$gamma2)

ts_data <- ts(data$gamma2, 
              start = c(2024), 
              frequency = 12)

# decompose time series
ts_decomposition <- decompose(ts_data)

plot(ts_decomposition)

# ARIMA model
arima_model <- auto.arima(ts_data, seasonal = TRUE)
summary(arima_model)

checkresiduals(arima_model)

# forecast next 12 months
forecast <- forecast(arima_model, h = 12)
summary(forecast)

plot(forecast, main = "Forecast for Next 12 Months", xlab = "Time", ylab = "gamma2")


```

```{r}


library(dplyr)
library(lubridate)  

daily_avg_2 <- data %>%
  drop_na(gamma2) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma2 = mean(gamma2, na.rm = TRUE))

plot.ts(daily_avg_2)

plot(daily_avg_2$date_only, daily_avg_2$avg_gamma2)

# gamma 3
daily_avg_3 <- data %>%
  drop_na(gamma3) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma3 = mean(gamma3, na.rm = TRUE))

plot.ts(daily_avg_3)

plot(daily_avg_3$date_only, daily_avg_3$avg_gamma3)

# gamma 4
daily_avg_4 <- data %>%
  drop_na(gamma4) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma4 = mean(gamma4, na.rm = TRUE))

plot.ts(daily_avg_4)

plot(daily_avg_4$date_only, daily_avg_4$avg_gamma4)


# gamma 5
daily_avg_5 <- data %>%
  drop_na(gamma5) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma5 = mean(gamma5, na.rm = TRUE))

plot.ts(daily_avg_5)

plot(daily_avg_5$date_only, daily_avg_5$avg_gamma5)


# gamma 6
daily_avg_6 <- data %>%
  drop_na(gamma6) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma6 = mean(gamma6, na.rm = TRUE))

plot.ts(daily_avg_6)

plot(daily_avg_6$date_only, daily_avg_6$avg_gamma6)


# gamma 7 
daily_avg_7 <- data %>%
  drop_na(gamma7) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma7 = mean(gamma7, na.rm = TRUE))

plot.ts(daily_avg_7)

plot(daily_avg_7$date_only, daily_avg_7$avg_gamma7)


# gamma 8
daily_avg_8 <- data %>%
  drop_na(gamma8) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma8 = mean(gamma8, na.rm = TRUE))

plot.ts(daily_avg_8)

plot(daily_avg_8$date_only, daily_avg_8$avg_gamma8)


# gamma 9
daily_avg_9 <- data %>%
  drop_na(gamma9) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_gamma9 = mean(gamma9, na.rm = TRUE))

plot.ts(daily_avg_9)

plot(daily_avg_9$date_only, daily_avg_9$avg_gamma9)


# dose
daily_avg_d <- data %>%
  drop_na(doseEquivalent) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_dose = mean(doseEquivalent, na.rm = TRUE))

plot.ts(daily_avg_d)

plot(daily_avg_d$date_only, daily_avg_d$avg_dose)



# gamma sum
daily_avg_sum <- data %>%
  drop_na(gammaSum) %>%
  mutate(date_only = as.Date(time)) %>%  # replace 'datetime_column' with your actual column name
  group_by(date_only) %>%
  summarize(avg_sum = mean(gammaSum, na.rm = TRUE))

plot.ts(daily_avg_sum)

plot(daily_avg_sum$date_only, daily_avg_sum$avg_sum)







```



ARIMA(p,d,q)(P,D,Q)[24] — to model daily seasonality



Daily time series
```{r}
library(dplyr)
library(lubridate)
library(forecast)

# Make sure timestamp column is POSIXct
# Replace `your_timestamp_col` with your actual datetime column name
data <- data %>%
  mutate(datetime = as.POSIXct(time)) %>%
  drop_na(gamma2)

# Aggregate to daily (mean or sum depending on your use case)
daily_data <- data %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date) %>%
  summarise(gamma2 = mean(gamma2))  # or sum(gamma2)

# Create time series object
ts_daily <- ts(daily_data$gamma2, frequency = 7)  # Weekly seasonality, can adjust if needed

# Decompose
ts_decomp <- stl(ts_daily, s.window = "periodic")
plot(ts_decomp)

# Fit ARIMA on daily data
arima_model <- auto.arima(ts_daily)

summary(arima_model)
checkresiduals(arima_model)

# Forecast for the rest of the year (e.g., next 365 days)
forecast_horizon <- 365
forecast_result <- forecast(arima_model, h = forecast_horizon)

plot(forecast_result, main = "Daily Forecast", xlab = "Day", ylab = "gamma2")


```
ARIMA(1,1,2)(1,0,2)[7]
non seasonal part (1,1,2)
AR = 1 autoregressive term
I first-order differencing, stationary
MA = 2 moving average terms
seasonal part (1,0,2)[7]
SAR = 1 seasonal autoregressive term
SI no seasonal difference
SMA 2 seasonal moving average terms
seasonality every 7 days

ME, forecasts slightly underestimating
RMSE heavily penalizes large errors
MAE average magnitude of the forecast errors
MAPE pretty good, errors are small relative to actuals
MASE <1 means its better than a naive forecast
ACF1 (autocorrelation at lag 1) very close to 0 so residuals are basically white noise which is good

Ljung-box test
- tests whether residuals are uncorrelated
- p-value not quite significant, so no strong evidence of autocorrelation in the residuals which is good
***what does Q represent???


forecast daily instead of hourly because takes long time to compute
```{r}
library(forecast)
library(lubridate)

gamma_column <- 'gammaSum'


# Convert gamma2 column to a daily time series (starting from 2008)
gamma_ts <- ts(data[[gamma_column]], frequency = 365, start = c(2008, 1))

# Fit Seasonal and Trend decomposition using LOESS + forecast
# STLF combines STL decomposition with an ARIMA or ETS forecast on the seasonally adjusted series
gamma_forecast_stlf <- stlf(gamma_ts, h = as.integer(as.Date("2025-03-26") - as.Date("2025-01-01")))

# Summary of the forecast (includes point forecast, confidence intervals, etc.)
summary(gamma_forecast_stlf)

# Create date sequence matching forecast
forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))

# Plot: forecast line only
plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

# Plot: with 95% confidence interval shaded
plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"),
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))

polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # shaded area for 95% CI

lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)

legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", rgb(1, 0, 0, 0.2)), lty = c(1, NA), lwd = c(2, NA), fill = c(NA, rgb(1, 0, 0, 0.2)))

plot.ts(lincoln25$`GAMMA COUNT RATE R02 (CPM)`)



```
STFL (Seasonal and Trend Decomposition)
- using LOESS, then forecasting

STL Decomposition, splits time series into trend, seasonal, remainder(noise), 
- good for daily data where seasonal patterns repeat

deseasonalized forecast
- remove seasonal component, remainder+trend is forecasted using ETS
- after forecast, seasonal component added back to get final forecast


***interpret summary of forecast
alpha=0.999
beta=0.3386
phi = 0.8
initial states?

is there not a p-value in the summary of forecast? or only in arima model


gamma 2 doesn't give very accurate prediction for beginning of 2025, dose was predicted well, but how do the other gamma channels predict



```{r}
# lincoln 25 outliers

no_outliers_25 <- subset(lincoln25, `GAMMA COUNT RATE R02 (CPM)` >1700 & `GAMMA COUNT RATE R02 (CPM)` <2000)

ts.plot(no_outliers_25$`GAMMA COUNT RATE R02 (CPM)`)


```


```{r}
un <- unique(lincoln25$`SAMPLE COLLECTION TIME`)

for (day in un) {
  day_av <- mean(lincoln25$`SAMPLE COLLECTION TIME`)
}


```

