---
title: "gamma"
output: pdf_document
date: "2025-03-31"
---
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(lubridate)
library(parallel)
library(forecast)

lincoln08 <- read_csv("NE_LINCOLN_2008.csv")
plot.ts(lincoln08$`GAMMA COUNT RATE R02 (CPM)`) # plot time series over that year

lincoln09 <- read_csv("NE_LINCOLN_2009.csv")
plot.ts(lincoln09$`GAMMA COUNT RATE R02 (CPM)`)

lincoln10 <- read_csv("NE_LINCOLN_2010.csv")
lincoln11 <- read_csv("NE_LINCOLN_2011.csv")
lincoln12 <- read_csv("NE_LINCOLN_2012.csv")
lincoln13 <- read_csv("NE_LINCOLN_2013.csv")
lincoln14 <- read_csv("NE_LINCOLN_2014.csv")
lincoln15 <- read_csv("NE_LINCOLN_2015.csv")
lincoln16 <- read_csv("NE_LINCOLN_2016.csv")
lincoln17 <- read_csv("NE_LINCOLN_2017.csv")
lincoln18 <- read_csv("NE_LINCOLN_2018.csv")
lincoln19 <- read_csv("NE_LINCOLN_2019.csv")
lincoln20 <- read_csv("NE_LINCOLN_2020.csv")
lincoln21 <- read_csv("NE_LINCOLN_2021.csv")
lincoln22 <- read_csv("NE_LINCOLN_2022.csv")
lincoln23 <- read_csv("NE_LINCOLN_2023.csv")
lincoln24 <- read_csv("NE_LINCOLN_2024.csv")
lincoln25 <- read_csv("NE_LINCOLN_2025 (1).csv")

```

Combine data
```{r}
all_data <- bind_rows(lincoln08, lincoln09, lincoln10, lincoln11, lincoln12,
                      lincoln13, lincoln14, lincoln15, lincoln16, lincoln17,
                      lincoln18, lincoln19, lincoln20, lincoln21, lincoln22,
                      lincoln23, lincoln24, lincoln25)
head(all_data)


plot.ts(all_data$`GAMMA COUNT RATE R02 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R02 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R03 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R03 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R04 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R04 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R05 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R05 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R06 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R06 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R07 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R07 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R08 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R08 (CPM)`)

plot.ts(all_data$`GAMMA COUNT RATE R09 (CPM)`)
summary(all_data$`GAMMA COUNT RATE R09 (CPM)`)

plot.ts(all_data$`DOSE EQUIVALENT RATE (nSv/h)`)
summary(all_data$`DOSE EQUIVALENT RATE (nSv/h)`)





```







R02
```{r}
# format collection time
all_data$SAMPLE_COLLECTION_TIME <- as.POSIXct(all_data$`SAMPLE COLLECTION TIME`, format="%m/%d/%Y %H:%M:%S")

gamma_column <- 'GAMMA COUNT RATE R02 (CPM)'

# convert to time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))


plot.ts(lincoln25$`GAMMA COUNT RATE R02 (CPM)`)


#######
```

summary(gamma_forecast_stlf) output

Method:
STL + ETS(M,Ad,N)  
- combination of seasonal and trend decomposition using Loess(STL) and Exponential Smoothing State Space Model (ETS)
- M= multiplicative trend, Ad=additive seasonality, N=no error adjustment

smoothign parameters
alpha=0.9988, high which means model closely tracks most recent observations
beta=0.3546, determines how much model respoinds to changes in trend
phi=0.8, damping parameter, affects how much trend value is reduced over time

Initial states:
I=2962.83, initial level
b=21.48, initial rate of change

sigma = 0.0394,

AIC=2577517
BIC=2577575

MSE=-0.0051, small bias in predictions, model slighly underpredicts on average
RMSE=89.016
MAE=47.434
MASE=0.1742, want value closer to 1 for good model
ACF1=0.1335, want value close to 0 meaning residuals aren't correlated

Summary
- low bias but high RMSE and MAE so might be considerable errors in forecast
- model captures most of underlying patters in the data




```{r}
# R03

gamma_column <- 'GAMMA COUNT RATE R03 (CPM)'

# convert to time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))


plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))

plot.ts(lincoln25$`GAMMA COUNT RATE R03 (CPM)`)

```

```{r}
# R04

gamma_column <- 'GAMMA COUNT RATE R04 (CPM)'

# convert to time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))

plot.ts(lincoln25$`GAMMA COUNT RATE R04 (CPM)`)

```

```{r}

# R05

gamma_column <- 'GAMMA COUNT RATE R05 (CPM)'

# convert to time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))


plot.ts(lincoln25$`GAMMA COUNT RATE R05 (CPM)`)

```

```{r}
#######

# R06

gamma_column <- 'GAMMA COUNT RATE R06 (CPM)'

# convert to time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))


plot.ts(lincoln25$`GAMMA COUNT RATE R06 (CPM)`)


```

```{r}

# R07

gamma_column <- 'GAMMA COUNT RATE R07 (CPM)'

# convert to  time series object (daily data)

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))


plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))


plot.ts(lincoln25$`GAMMA COUNT RATE R07 (CPM)`)

```

```{r}
#######

# R08

gamma_column <- 'GAMMA COUNT RATE R08 (CPM)'

# convert to a time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))


plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))

plot.ts(lincoln25$`GAMMA COUNT RATE R08 (CPM)`)


```

```{r}

# R09

gamma_column <- 'GAMMA COUNT RATE R09 (CPM)'

# convert to  time series object (daily data)
gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

# fit the Seasonal and Trend decomposition using LOESS (stlf) model on the historical data (2008-2025)
gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-03-26") - as.Date("2025-01-01"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
summary(gamma_forecast_stlf)

forecast_dates <- seq(as.Date("2025-01-01"), by = "day", length.out = length(gamma_forecast_stlf$mean))


# plot forecast for first three months, plot actual 2025 data and

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))


plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"), 
     ylim = range(c(gamma_forecast_stlf$lower[,2], gamma_forecast_stlf$upper[,2], gamma_forecast_stlf$mean)))
polygon(c(forecast_dates, rev(forecast_dates)), 
        c(gamma_forecast_stlf$lower[,2], rev(gamma_forecast_stlf$upper[,2])), 
        col = rgb(1, 0, 0, 0.2), border = NA)  # Light red shaded area
lines(forecast_dates, gamma_forecast_stlf$mean, col = "blue", lwd = 2)
legend("topleft", legend = c("Forecast", "95% Confidence Interval"),
       col = c("blue", "pink"), lty = c(1, 1), lwd = c(2, 5), fill = c(NA, rgb(1, 0, 0, 0.2)))


plot.ts(lincoln25$`GAMMA COUNT RATE R09 (CPM)`)

```

```{r}
# only dealing with data after 3/26 each year
data_2024 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2024-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2025-01-01"))

data_2023 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2023-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2024-01-01"))
data_2022 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2022-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2023-01-01"))
data_2021 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2021-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2022-01-01"))
data_2020 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2020-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2021-01-01"))
data_2019 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2019-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2020-01-01"))
data_2018 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2018-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2019-01-01"))
data_2017 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2017-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2018-01-01"))
data_2016 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2016-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2017-01-01"))
data_2015 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2015-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2016-01-01"))
data_2014 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2014-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2015-01-01"))
data_2013 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2013-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2014-01-01"))
data_2012 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2012-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2013-01-01"))
data_2011 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2011-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2012-01-01"))
data_2010 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2010-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2011-01-01"))
data_2009 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2009-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2010-01-01"))
data_2008 <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2008-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2009-01-01"))


plot(data_2008$SAMPLE_COLLECTION_TIME, data_2008$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2008")
plot(data_2009$SAMPLE_COLLECTION_TIME, data_2009$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2009")
plot(data_2010$SAMPLE_COLLECTION_TIME, data_2010$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2010")
plot(data_2011$SAMPLE_COLLECTION_TIME, data_2011$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2011")
plot(data_2012$SAMPLE_COLLECTION_TIME, data_2012$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2012")


par(mfrow = c(2,3))
plot(data_2013$SAMPLE_COLLECTION_TIME, data_2013$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2013")
plot(data_2014$SAMPLE_COLLECTION_TIME, data_2014$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2014")
plot(data_2015$SAMPLE_COLLECTION_TIME, data_2015$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2015")
plot(data_2016$SAMPLE_COLLECTION_TIME, data_2016$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2016")
plot(data_2017$SAMPLE_COLLECTION_TIME, data_2017$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2017")
plot(data_2018$SAMPLE_COLLECTION_TIME, data_2018$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2018")


par(mfrow = c(2,3))
plot(data_2019$SAMPLE_COLLECTION_TIME, data_2019$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2019")
plot(data_2020$SAMPLE_COLLECTION_TIME, data_2020$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2020")
plot(data_2021$SAMPLE_COLLECTION_TIME, data_2021$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2021")
plot(data_2022$SAMPLE_COLLECTION_TIME, data_2022$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2022")
plot(data_2023$SAMPLE_COLLECTION_TIME, data_2023$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2023")
# Create an empty plot for the year 2024 data
plot(data_2024$SAMPLE_COLLECTION_TIME, data_2024$`GAMMA COUNT RATE R02 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R02 (CPM)", 
     main = "R02 for 2024")



```


```{r}
all_data # make sure has all years
# convert collection time
all_data$SAMPLE_COLLECTION_TIME <- as.POSIXct(all_data$`SAMPLE COLLECTION TIME`, format="%m/%d/%Y %H:%M:%S")

# filter data for all years (2008 to 2024), and only for the months April to December
data_filtered <- all_data %>%
  filter(SAMPLE_COLLECTION_TIME >= as.POSIXct("2008-03-27") & SAMPLE_COLLECTION_TIME < as.POSIXct("2025-01-01")) %>%
  mutate(Month = month(SAMPLE_COLLECTION_TIME), Year = year(SAMPLE_COLLECTION_TIME)) %>%
  filter(Month >= 4 & Month <= 12)  # Only keep data from April to December

plot(data_filtered$Month,data_filtered$`GAMMA COUNT RATE R02 (CPM)`)
summary(data_filtered)

```




R03
```{r}
gamma_column <- 'GAMMA COUNT RATE R03 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))


par(mfrow = c(3,3))
plot(data_2008$SAMPLE_COLLECTION_TIME, data_2008$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2008")
plot(data_2009$SAMPLE_COLLECTION_TIME, data_2009$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2009")
plot(data_2010$SAMPLE_COLLECTION_TIME, data_2010$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2010")
plot(data_2011$SAMPLE_COLLECTION_TIME, data_2011$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2011")
plot(data_2012$SAMPLE_COLLECTION_TIME, data_2012$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2012")
plot(data_2013$SAMPLE_COLLECTION_TIME, data_2013$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2013")
plot(data_2014$SAMPLE_COLLECTION_TIME, data_2014$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2014")
plot(data_2015$SAMPLE_COLLECTION_TIME, data_2015$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2015")
plot(data_2016$SAMPLE_COLLECTION_TIME, data_2016$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2016")

par(mfrow = c(2,4))
plot(data_2017$SAMPLE_COLLECTION_TIME, data_2017$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2017")
plot(data_2018$SAMPLE_COLLECTION_TIME, data_2018$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2018")
plot(data_2019$SAMPLE_COLLECTION_TIME, data_2019$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2019")
plot(data_2020$SAMPLE_COLLECTION_TIME, data_2020$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2020")
plot(data_2021$SAMPLE_COLLECTION_TIME, data_2021$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2021")
plot(data_2022$SAMPLE_COLLECTION_TIME, data_2022$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2022")
plot(data_2023$SAMPLE_COLLECTION_TIME, data_2023$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2023")
# Create an empty plot for the year 2024 data
plot(data_2024$SAMPLE_COLLECTION_TIME, data_2024$`GAMMA COUNT RATE R03 (CPM)`, 
     type = "l", col = "blue", 
     xlab = "Date", ylab = "R03 (CPM)", 
     main = "R03 for 2024")


```


R04
```{r}
gamma_column <- 'GAMMA COUNT RATE R04 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

```

R05
```{r}
gamma_column <- 'GAMMA COUNT RATE R05 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

```

R06
```{r}
gamma_column <- 'GAMMA COUNT RATE R06 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

```

R07
```{r}
gamma_column <- 'GAMMA COUNT RATE R07 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

```

R08
```{r}
gamma_column <- 'GAMMA COUNT RATE R08 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

```


R09
```{r}
gamma_column <- 'GAMMA COUNT RATE R09 (CPM)'

gamma_ts <- ts(all_data[[gamma_column]], frequency = 365, start = c(2008, 1))

gamma_model_stlf <- stlf(gamma_ts)

forecast_days <- as.integer(as.Date("2025-12-31") - as.Date("2025-03-27"))

gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)

forecast_dates <- seq(as.Date("2025-03-27"), by = "day", length.out = length(gamma_forecast_stlf$mean))

plot(forecast_dates, gamma_forecast_stlf$mean, type = "l", col = "blue", 
     xlab = "Date", ylab = paste("Forecast for", gamma_column), 
     main = paste("Forecast for", gamma_column, "in 2025"))

```


```{r}
gamma_columns <- c('GAMMA COUNT RATE R02 (CPM)', 'GAMMA COUNT RATE R03 (CPM)', 
                   'GAMMA COUNT RATE R04 (CPM)', 'GAMMA COUNT RATE R05 (CPM)', 
                   'GAMMA COUNT RATE R06 (CPM)', 'GAMMA COUNT RATE R07 (CPM)', 
                   'GAMMA COUNT RATE R08 (CPM)', 'GAMMA COUNT RATE R09 (CPM)', 
                   'DOSE EQUIVALENT RATE (nSv/h)')


forecast_mean_values <- list()

for (i in seq_along(gamma_columns)) {
  gamma_ts <- ts(all_data[[gamma_columns[i]]], frequency = 365, start = c(2008, 1))

  gamma_model_stlf <- stlf(gamma_ts)

  forecast_days <- length(forecast_dates)
  gamma_forecast_stlf <- forecast(gamma_model_stlf, h = forecast_days)
  
  forecast_mean_values[[i]] <- gamma_forecast_stlf$mean
}

all_forecasts <- unlist(forecast_mean_values)

y_limits <- range(all_forecasts, na.rm = TRUE) 

plot(forecast_dates, rep(NA, length(forecast_dates)), type = "n", 
     xlab = "Date", ylab = "Gamma Count Rate (CPM)", 
     main = "Forecast for Gamma Count Rates in 2025", 
     ylim = y_limits)  # Set y-axis limits based on the forecast data

colors <- c("blue", "green", "red", "purple", "orange", "brown", "pink", "cyan", 'black')

for (i in seq_along(gamma_columns)) {
  lines(forecast_dates, forecast_mean_values[[i]], col = colors[i], lwd = 2)
}
legend("topright", legend = gamma_columns, col = colors, lwd = 2, bty = "n", cex = 0.8)


```
